
#Ligand-receptor analysis for Somatic cells in Mouse (using ligand/receptor interaction data from Ramilowski_2015)

567 Receptors
586 Ligands
2015 Interactions

##1. Select highly variable genes from somatic and germline clusters clusters using Jun's code
```{r}
soma.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/MouseAdultST25_7SomaticCellTypes_Centroids_Qianyi.txt",header = T,sep="\t",row.names = 1))
colnames(soma.mean) <- c("InnateLymphoid", "Macrophage", "Endothelial", "Myoid", "Leydig","Sertoli","Tcf21+")
spg.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/SPG_centroids/MouseSPGCluster_AllGene_Centroids.txt",header = T,sep="\t",row.names = 1))
```


## function used for plotting dots with density as heat
```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
library(viridis)
library(ggExtra)
library(gridExtra)
```

###. Select highly variable genes across stages.
```{r}
s.mean <- apply(soma.mean,1,mean)
s.var <- apply(soma.mean,1,var)

s.var <- s.var[s.mean!=0]
s.mean <- s.mean[s.mean!=0]
```

```{r}
filter.s <- names(s.mean)[s.mean > 0.15 & s.var/s.mean > 0.15]
length(filter.s)
soma.gen <- data.frame(mean = s.mean, varmean = s.var/s.mean)
soma.gen$den <- get_density(soma.gen$mean, soma.gen$varmean, n=100)

p <- ggplot(soma.gen) + geom_point(aes(x=mean, y=varmean, color = log(den+1)),size=0.5) + scale_color_viridis()+ ggtitle("Soma HVGs") + theme(legend.position="none") + xlab("Mean") + ylab("Variation/Mean") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + geom_hline(yintercept = 0.15) + geom_vline(xintercept = 0.15)
p2 <- ggMarginal(p, type="density")
grid.arrange(p2, ncol=1)

```

```{r}
somaGenes <- filter.s
```

##2. Read in Ligand and Receptor data
```{r}
lrinter.hum <- read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/Ramilowski_LigRec_Human.txt",header = TRUE)
```
###. Human-Mouse
```{r}
library(biomaRt) #Ensemble Gene 94
ensembl <- useMart(host='http://oct2018.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL')
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name")

humMou <- getBM(attributes = attributes,mart = ensembl)
dim(humMou)

#Remove non-homologs
humMouOrth <- humMou[humMou$mmusculus_homolog_ensembl_gene!='',]
dim(humMouOrth)  ##26312
```


```{r}
lrinter <- c()
for(i in c(1:dim(lrinter.hum)[1])){
  lh <- as.character(lrinter.hum[i,1])
  rh <- as.character(lrinter.hum[i,2])
  if(!lh %in% humMouOrth$external_gene_name){
    next
  }
  if(!rh %in% humMouOrth$external_gene_name){
    next
  }  
  lms <- humMouOrth[humMouOrth$external_gene_name==lh,]$mmusculus_homolog_associated_gene_name
  rms <- humMouOrth[humMouOrth$external_gene_name==rh,]$mmusculus_homolog_associated_gene_name
  for(i in lms){
    for(j in rms){
      lrinter <- rbind(lrinter, c(i,j))
    }
  }
}
```

```{r}
ligands <- unique(lrinter[,1])
receptors <- unique(lrinter[,2])
```
##3. Extract Ligands and Receptors from Germline and Sertoli cells
```{r}
soma.ligands <- intersect(somaGenes,ligands)
soma.receptors <- intersect(somaGenes,receptors)
length(soma.ligands) #Number of ligands in somatic cells
## [1] 142
length(soma.receptors) #Number of receptors in somatic cells
## [1] 166
```


##4. Find ligand receptor pairs from highly variable genes
```{r}
ligrec <- function(ligands,receptors){
  lig.rec.pairs <- c()
  for(i in c(1:dim(lrinter)[1])){
    if(lrinter[i,1] %in% ligands | lrinter[i,2] %in% receptors){
      if(lrinter[i,1] %in% rownames(soma.mean) & lrinter[i,2] %in% rownames(soma.mean)){
        lig.rec.pairs <- rbind(lig.rec.pairs, c(as.character(lrinter[i,1]), as.character(lrinter[i,2])))
      }
  }
  }
  return(lig.rec.pairs)
}
lig.rec.soma.soma <- ligrec(soma.ligands,soma.receptors)
dim(lig.rec.soma.soma)[1] # Number of pairs for Ligand(Soma)-Receptor(Germ)
# 1011
```

```{r}
soma.lab <- colnames(soma.mean)
```

```{r}
soma.combs <- c()
for(i in soma.lab){
  for(j in soma.lab){
    comb <- paste(i, j, sep = "-")
    soma.combs <- c(soma.combs, comb)
  }
}
int.soma.mat <- matrix(0, nrow = dim(lig.rec.soma.soma)[1], ncol = length(soma.combs))
rownames(int.soma.mat) <- apply(lig.rec.soma.soma,1,function(x) paste(x[1],x[2],sep = "-"))
colnames(int.soma.mat) <- soma.combs

for(i in soma.lab){
  cat(i)
  for(j in soma.lab){
    for(k in c(1:dim(lig.rec.soma.soma)[1])){
        lig <- as.character(lig.rec.soma.soma[k,1])
        rec <- as.character(lig.rec.soma.soma[k,2])
        int.pairs <- paste(lig,rec,sep = "-")
        cel.pairs <- paste(i,j,sep = "-")
        int.scores <- soma.mean[lig,i] * soma.mean[rec,j]
        int.soma.mat[int.pairs,cel.pairs] <- int.scores
    }
  }
}

```

```{r}
saveRDS(int.soma.mat, file = "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200217_Soma_Soma/ehvg.soma.mou.int.soma.mat.rds")
```

