
#Ligand-receptor analysis for SPG cells and Somatic cells in Mouse (using ligand/receptor interaction data from Ramilowski_2015)

567 Receptors
586 Ligands
2015 Interactions

##1. Select highly variable genes from somatic and germline clusters clusters using Jun's code
```{r}
soma.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/MouseAdultST25_7SomaticCellTypes_Centroids_Qianyi.txt",header = T,sep="\t",row.names = 1))
spg.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/SPG_centroids/MouseSPGCluster_AllGene_Centroids.txt",header = T,sep="\t",row.names = 1))
germ.20.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/mou.20.warpped.centroids.text", skip = 1, row.names = 1, sep = "\t"))
germ.mean <- cbind(spg.mean, germ.20.mean)
germ.mean <- log(germ.mean+1)
##data source: https://umich.app.box.com/folder/31463886658
##First 1-Mar and 2-Mar are removed because of duplicates.
```


## function used for plotting dots with density as heat
```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
library(viridis)
library(ggExtra)
library(gridExtra)
```

###. Select highly variable genes across stages.
```{r}
g.mean<-apply(germ.mean,1,mean)
g.var<-apply(germ.mean,1,var)

s.mean <- apply(soma.mean,1,mean)
s.var <- apply(soma.mean,1,var)

g.var <- g.var[g.mean!=0]
g.mean <- g.mean[g.mean!=0]
s.var <- s.var[s.mean!=0]
s.mean <- s.mean[s.mean!=0]
```

```{r}
filter.g <- names(g.mean)[g.mean > 0.08 & g.var/g.mean > 0.08]
length(filter.g)
germ.gen <- data.frame(mean = g.mean, varmean = g.var/g.mean)
germ.gen$den <- get_density(germ.gen$mean, germ.gen$varmean, n=100)

p <- ggplot(germ.gen) + geom_point(aes(x=mean, y=varmean, color = log(den+1)),size=0.5) + scale_color_viridis()+ ggtitle("Germ HVGs") + theme(legend.position="none") + xlab("Mean") + ylab("Variation/Mean") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + geom_hline(yintercept = 0.08) + geom_vline(xintercept = 0.08)
p1 <- ggMarginal(p, type="density")
grid.arrange(p1, ncol=1)

filter.s <- names(s.mean)[s.mean > 0.15 & s.var/s.mean > 0.15]
length(filter.s)
soma.gen <- data.frame(mean = s.mean, varmean = s.var/s.mean)
soma.gen$den <- get_density(soma.gen$mean, soma.gen$varmean, n=100)

p <- ggplot(soma.gen) + geom_point(aes(x=mean, y=varmean, color = log(den+1)),size=0.5) + scale_color_viridis()+ ggtitle("Soma HVGs") + theme(legend.position="none") + xlab("Mean") + ylab("Variation/Mean") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + geom_hline(yintercept = 0.15) + geom_vline(xintercept = 0.15)
p2 <- ggMarginal(p, type="density")
grid.arrange(p2, ncol=1)

```

```{r}
somaGenes <- filter.s
germGenes <- filter.g
```

##2. Read in Ligand and Receptor data
```{r}
lrinter.hum <- read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/Ramilowski_LigRec_Human.txt",header = TRUE)
```
###. Human-Mouse
```{r}
library(biomaRt) #Ensemble Gene 94
ensembl <- useMart(host='http://oct2018.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL')
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name")

humMou <- getBM(attributes = attributes,mart = ensembl)
dim(humMou)

#Remove non-homologs
humMouOrth <- humMou[humMou$mmusculus_homolog_ensembl_gene!='',]
dim(humMouOrth)  ##26312
```


```{r}
lrinter <- c()
for(i in c(1:dim(lrinter.hum)[1])){
  lh <- as.character(lrinter.hum[i,1])
  rh <- as.character(lrinter.hum[i,2])
  if(!lh %in% humMouOrth$external_gene_name){
    next
  }
  if(!rh %in% humMouOrth$external_gene_name){
    next
  }  
  lms <- humMouOrth[humMouOrth$external_gene_name==lh,]$mmusculus_homolog_associated_gene_name
  rms <- humMouOrth[humMouOrth$external_gene_name==rh,]$mmusculus_homolog_associated_gene_name
  for(i in lms){
    for(j in rms){
      lrinter <- rbind(lrinter, c(i,j))
    }
  }
}
```

```{r}
ligands <- unique(lrinter[,1])
receptors <- unique(lrinter[,2])
```
##3. Extract Ligands and Receptors from Germline and Sertoli cells
```{r}
germ.ligands <- intersect(germGenes,ligands)
soma.receptors <- intersect(somaGenes,receptors)
length(germ.ligands) #Number of ligands in germline cells
## [1] 71
length(soma.receptors) #Number of receptors in somatic cells
## [1] 166
```


##4. Find ligand receptor pairs from highly variable genes
```{r}
ligrec <- function(ligands,receptors){
  lig.rec.pairs <- c()
  for(i in c(1:dim(lrinter)[1])){
    if(lrinter[i,1] %in% ligands | lrinter[i,2] %in% receptors){
      if(lrinter[i,1] %in% rownames(germ.mean) & lrinter[i,2] %in% rownames(soma.mean)){
        lig.rec.pairs <- rbind(lig.rec.pairs, c(as.character(lrinter[i,1]), as.character(lrinter[i,2])))
      }
  }
  }
  return(lig.rec.pairs)
}
lig.rec.germ.soma <- ligrec(germ.ligands,soma.receptors)
dim(lig.rec.germ.soma)[1] # Number of pairs for Ligand(Soma)-Receptor(Germ)
# 1039
```

```{r}
write.table(lig.rec.germ.soma, file = "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200220_Germ_Soma/eitherHVG.mou.lig.rec.germ.soma.txt", quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t")
```


```{r}
germ.lab <- c(paste0("SPG",c(1:6)), paste0("GC",c(1:20)))
colnames(germ.mean) <- germ.lab
soma.lab <- colnames(soma.mean)
```

```{r}
cells.combs <- c()
for(i in germ.lab){
  for(j in soma.lab){
    comb <- paste(i, j, sep = "-")
    cells.combs <- c(cells.combs, comb)
  }
}
soma.germ.int.cells.mat <- matrix(0, nrow = dim(lig.rec.germ.soma)[1], ncol = length(cells.combs))
rownames(soma.germ.int.cells.mat) <- apply(lig.rec.germ.soma,1,function(x) paste(x[1],x[2],sep = "-"))
colnames(soma.germ.int.cells.mat) <- cells.combs

for(i in germ.lab){
  cat(i)
  for(j in soma.lab){
    for(k in c(1:dim(lig.rec.germ.soma)[1])){
        lig <- as.character(lig.rec.germ.soma[k,1])
        rec <- as.character(lig.rec.germ.soma[k,2])
        int.pairs <- paste(lig,rec,sep = "-")
        cel.pairs <- paste(i,j,sep = "-")
        int.scores <- germ.mean[lig,i] * soma.mean[rec,j]
        soma.germ.int.cells.mat[int.pairs,cel.pairs] <- int.scores
    }
  }
}

```

```{r}
saveRDS(soma.germ.int.cells.mat, file = "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200220_Germ_Soma/ehvg.germ.soma.int.cells.mat.rds")
```


```{r}
all.scores <- as.vector(soma.germ.int.cells.mat)
quant.95 <- quantile(all.scores, probs = 0.95)
soma.germ.int.cells.mat.trc <- soma.germ.int.cells.mat
soma.germ.int.cells.mat.trc[soma.germ.int.cells.mat<quant.95] = 0
```

```{r}
int.mean <- apply(soma.germ.int.cells.mat.trc,1,mean)
soma.germ.int.cells.mat.trc <- soma.germ.int.cells.mat.trc[int.mean!=0, ]
```

```{r}
soma.germ.mou.int <- apply(soma.germ.int.cells.mat.trc,2,function(x) sum(x>0))
```

```{r}
#soma.germ.mou.int[grepl("SPG",names(soma.germ.mou.int))]=0
```


```{r}
soma.germ.mou.df <- data.frame(Interactions = soma.germ.mou.int,Ligand=unlist(lapply(names(soma.germ.mou.int),function(x) str_split(x,"-")[[1]][1])), Receptor=unlist(lapply(names(soma.germ.mou.int),function(x) str_split(x,"-")[[1]][2])))
levels(soma.germ.mou.df$Receptor)[levels(soma.germ.mou.df$Receptor)=="Unknown"]="Tcf21+Interstitial"
soma.germ.mou.df$Receptor <- factor(soma.germ.mou.df$Receptor, levels = c("InnateLymphoid","Macrophage","Endothelial","Myoid","Sertoli","Leydig","Tcf21+Interstitial"))
soma.germ.mou.df$Ligand <- factor(soma.germ.mou.df$Ligand, levels = c(paste0("SPG",c(1:6)), paste0("GC",c(1:20))))
mou.lig.total = soma.germ.mou.df %>% group_by(Ligand) %>% summarise(sum=sum(Interactions))
mou.rec.total = soma.germ.mou.df %>% group_by(Receptor) %>% summarise(sum=sum(Interactions))
```

```{r fig.width=6, fig.height=6}
int.circ.plot2 <- function(lig.rec, rad1, rad2, lig.lab, rec.lab, lig.col, rec.col,lig.label, rec.label, lig.total, rec.total){
#pdf(filename,width = 6*lineup[2],height = 6*lineup[1])

lig.len = length(lig.lab)
rec.len = length(rec.lab)
plot(1, type="n", xlab="", ylab="", xlim=c(-2, 17), ylim=c(0, max(lig.len,rec.len)+2),axes = FALSE) ##axes2 = FALSE

for(i in c(1:lig.len)){
  if(lig.len >= rec.len){
    h = i
  }else{
    h = 2.5*i + ceiling((rec.len-2.5*lig.len)/2)
  }
  x_add <- sqrt(rad1^2-(12.5-(h-1))^2) - sqrt(rad1^2-12.5^2)
  draw.circle(3.5-x_add,h,rev(lig.total$sum)[i]/max(lig.total$sum)*0.5,border = rev(lig.col)[i],col = rev(lig.col)[i])
  text(2.8-x_add,h,labels = rev(lig.lab)[i],cex=1.2, adj=c(1,0.5))
}

#text(3.5,max(lig.len,rec.len)+2.8,paste0(lig.label,"\n(ligand)"),cex=1.5)
for(i in c(1:rec.len)){
  if(lig.len <= rec.len){
    h = i
  }else{
    h = 2.5*i + ceiling((lig.len-2.5*rec.len)/2)
  }
  x_add <- sqrt(rad2^2-(12.5-(h-1))^2) - sqrt(rad2^2-12.5^2)
  draw.circle(10.2+x_add,h,rev(rec.total$sum)[i]/max(rec.total$sum)*0.5,border = rev(rec.col)[i],col = rev(rec.col)[i])
  text(10.8+x_add,h,label = rev(rec.lab)[i],cex=1.2, adj=c(0,0.5))
}

#text(10.2,max(lig.len,rec.len)+2.8,paste0(rec.label,"\n(receptor)"),cex = 1.5)
shrink = 0.95
for(i in c(1:dim(lig.rec)[1])){
  if(lig.rec$Interactions[i]==0){
    next
  }
  if(lig.len >= rec.len){
    h1 = lig.len + 1
    h2 = 2.5*rec.len + 2.5 + ceiling((lig.len - 2.5*rec.len)/2)
    #h2 = lig.len + 1 - floor((lig.len - rec.len)/2)
  }else{
    #h2 = rec.len + 1
    h1 = 2.5*lig.len + 2.5 + ceiling((rec.len - 2.5*lig.len)/2)
    h2 = rec.len + 1
    #h2 = rec.len + 1
    #h1 = rec.len + 1
  }
  #y_height = h2-as.numeric(lig.rec$Receptor[i])
  left_heady <- h1-as.numeric(lig.rec$Ligand[i])
  left_x_add <- sqrt(rad1^2-(12.5-(left_heady-1))^2) - sqrt(rad1^2-12.5^2)
  headx = 3.3+shrink*(7-0.6)
  heady = h2-2.5*as.numeric(lig.rec$Receptor[i]) - (1-shrink)*((h2-2.5*as.numeric(lig.rec$Receptor[i])) - (h1-as.numeric(lig.rec$Ligand[i])))
  right_x_add <- sqrt(rad2^2-(12.5-(heady-1))^2) - sqrt(rad2^2-12.5^2)
  Arrows(3.8-left_x_add,left_heady,headx+right_x_add,heady,lwd = lig.rec$Interactions[i]/max(lig.rec$Interactions)*6, arr.width = 0.05, col = alpha("black", alpha = 0.6*(lig.rec$Interactions[i]-min(lig.rec$Interactions))/(max(lig.rec$Interactions)-min(lig.rec$Interactions))))
}
}
```

```{r fig.height=8, fig.width=7}
lig.lab <- as.character(unique(soma.germ.mou.df$Ligand))
rec.lab <- as.character(unique(soma.germ.mou.df$Receptor))
soma.col <- c("#006c44","#ffff33", "#984ea3", "#377eb8","#f781bf","#696969","#f0027f")
germ.col <- c(colorRampPalette(c("#c2d3fb", "#3269f3"))(6),colorRampPalette(c("#c2effb","#086b85"))(20))
#pdf(file = "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200204_Soma_Germ/Global.Mouse.inter.rotated.pdf", height = 8, width = 7)
int.circ.plot2(soma.germ.mou.df, 40, 40, lig.lab, rec.lab, germ.col, soma.col,"Germ","Soma",mou.lig.total, mou.rec.total)
#dev.off()
```

extract SPG-Tcf21+
```{r}
spg.tcf21 <- soma.germ.int.cells.mat.trc[,grepl("SPG.*Unknown",colnames(soma.germ.int.cells.mat.trc))]
spg.tcf21 <- spg.tcf21[apply(spg.tcf21,1,sum)!=0,]
colnames(spg.tcf21) <- paste0("SPG",c(1:6),"-Tcf21+")
```

```{r}
write.table(spg.tcf21, "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200220_Germ_Soma//Mouse.spg.tcf21.txt", row.names = TRUE, col.names = TRUE, sep = "\t", quote = FALSE)
```

