
#Ligand-receptor analysis for SPG cells and Somatic cells in Mouse (using ligand/receptor interaction data from Ramilowski_2015)

567 Receptors
586 Ligands
2015 Interactions

##1. Select highly variable genes from somatic and germline clusters clusters using Jun's code
```{r}
soma.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/MouseAdultST25_7SomaticCellTypes_Centroids_Qianyi.txt",header = T,sep="\t",row.names = 1))
spg.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/SPG_centroids/MouseSPGCluster_AllGene_Centroids.txt",header = T,sep="\t",row.names = 1))
germ.20.mean <- as.matrix(read.table("~/storage/HumanMacaqueMouseCompAnaysis/interdata/mou.20.warpped.centroids.text", skip = 1, row.names = 1, sep = "\t"))
germ.mean <- cbind(spg.mean, germ.20.mean)
germ.mean <- log(germ.mean+1)
##data source: https://umich.app.box.com/folder/31463886658
##First 1-Mar and 2-Mar are removed because of duplicates.
```


## function used for plotting dots with density as heat
```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
library(viridis)
library(ggExtra)
library(gridExtra)
```

###. Select highly variable genes across stages.
```{r}
g.mean<-apply(germ.mean,1,mean)
g.var<-apply(germ.mean,1,var)

s.mean <- apply(soma.mean,1,mean)
s.var <- apply(soma.mean,1,var)

g.var <- g.var[g.mean!=0]
g.mean <- g.mean[g.mean!=0]
s.var <- s.var[s.mean!=0]
s.mean <- s.mean[s.mean!=0]
```

```{r}
filter.g <- names(g.mean)[g.mean > 0.08 & g.var/g.mean > 0.08]
length(filter.g)
germ.gen <- data.frame(mean = g.mean, varmean = g.var/g.mean)
germ.gen$den <- get_density(germ.gen$mean, germ.gen$varmean, n=100)

p <- ggplot(germ.gen) + geom_point(aes(x=mean, y=varmean, color = log(den+1)),size=0.5) + scale_color_viridis()+ ggtitle("Germ HVGs") + theme(legend.position="none") + xlab("Mean") + ylab("Variation/Mean") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + geom_hline(yintercept = 0.08) + geom_vline(xintercept = 0.08)
p1 <- ggMarginal(p, type="density")
grid.arrange(p1, ncol=1)

filter.s <- names(s.mean)[s.mean > 0.15 & s.var/s.mean > 0.15]
length(filter.s)
soma.gen <- data.frame(mean = s.mean, varmean = s.var/s.mean)
soma.gen$den <- get_density(soma.gen$mean, soma.gen$varmean, n=100)

p <- ggplot(soma.gen) + geom_point(aes(x=mean, y=varmean, color = log(den+1)),size=0.5) + scale_color_viridis()+ ggtitle("Soma HVGs") + theme(legend.position="none") + xlab("Mean") + ylab("Variation/Mean") + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + geom_hline(yintercept = 0.15) + geom_vline(xintercept = 0.15)
p2 <- ggMarginal(p, type="density")
grid.arrange(p2, ncol=1)

```

```{r}
somaGenes <- filter.s
germGenes <- filter.g
```

##2. Read in Ligand and Receptor data
```{r}
lrinter.hum <- read.table("~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/Ramilowski_LigRec_Human.txt",header = TRUE)
```
###. Human-Mouse
```{r}
library(biomaRt) #Ensemble Gene 94
ensembl <- useMart(host='http://oct2018.archive.ensembl.org', 
                     biomart='ENSEMBL_MART_ENSEMBL')
ensembl = useDataset("hsapiens_gene_ensembl",mart=ensembl)
attributes = c("ensembl_gene_id","external_gene_name","mmusculus_homolog_ensembl_gene","mmusculus_homolog_associated_gene_name")

humMou <- getBM(attributes = attributes,mart = ensembl)
dim(humMou)

#Remove non-homologs
humMouOrth <- humMou[humMou$mmusculus_homolog_ensembl_gene!='',]
dim(humMouOrth)  ##26312
```


```{r}
lrinter <- c()
for(i in c(1:dim(lrinter.hum)[1])){
  lh <- as.character(lrinter.hum[i,1])
  rh <- as.character(lrinter.hum[i,2])
  if(!lh %in% humMouOrth$external_gene_name){
    next
  }
  if(!rh %in% humMouOrth$external_gene_name){
    next
  }  
  lms <- humMouOrth[humMouOrth$external_gene_name==lh,]$mmusculus_homolog_associated_gene_name
  rms <- humMouOrth[humMouOrth$external_gene_name==rh,]$mmusculus_homolog_associated_gene_name
  for(i in lms){
    for(j in rms){
      lrinter <- rbind(lrinter, c(i,j))
    }
  }
}
```

```{r}
ligands <- unique(lrinter[,1])
receptors <- unique(lrinter[,2])
```
##3. Extract Ligands and Receptors from Germline and Sertoli cells
```{r}
germ.ligands <- intersect(germGenes,ligands)
germ.receptors <- intersect(germGenes,receptors)
soma.ligands <- intersect(somaGenes,ligands)
soma.receptors <- intersect(somaGenes,receptors)
length(germ.ligands) #Number of ligands in germline cells
## [1] 71
length(germ.receptors) #Number of receptors in germline cells
## [1] 74
length(soma.ligands) #Number of ligands in somatic cells
## [1] 142
length(soma.receptors) #Number of receptors in somatic cells
## [1] 166
```


##4. Find ligand receptor pairs from highly variable genes
```{r}
ligrec <- function(ligands,receptors){
  lig.rec.pairs <- c()
  for(i in c(1:dim(lrinter)[1])){
    if(lrinter[i,1] %in% ligands | lrinter[i,2] %in% receptors){
      if(lrinter[i,1] %in% rownames(soma.mean) & lrinter[i,2] %in% rownames(germ.mean)){
        lig.rec.pairs <- rbind(lig.rec.pairs, c(as.character(lrinter[i,1]), as.character(lrinter[i,2])))
      }
  }
  }
  return(lig.rec.pairs)
}
lig.rec.soma.germ <- ligrec(soma.ligands,germ.receptors)
dim(lig.rec.soma.germ)[1] # Number of pairs for Ligand(Soma)-Receptor(Germ)
# 785
```

```{r}
write.table(lig.rec.soma.germ, file = "~/storage/HumanMacaqueMouseCompAnaysis/ligandReceptor/20200204_Soma_Germ/eitherHVG.mou.lig.rec.soma.germ.txt", quote = FALSE, row.names = FALSE, col.names = FALSE, sep = "\t")
```

```{r}
germ.lab <- c(paste0("SPG",c(1:6)), paste0("GC",c(1:20)))
colnames(germ.mean) <- germ.lab
soma.lab <- colnames(soma.mean)
```

```{r}
cells.combs <- c()
for(i in soma.lab){
  for(j in germ.lab){
    comb <- paste(i, j, sep = "-")
    cells.combs <- c(cells.combs, comb)
  }
}
int.cells.mat <- matrix(0, nrow = dim(lig.rec.soma.germ)[1], ncol = length(cells.combs))
rownames(int.cells.mat) <- apply(lig.rec.soma.germ,1,function(x) paste(x[1],x[2],sep = "-"))
colnames(int.cells.mat) <- cells.combs

for(i in soma.lab){
  cat(i)
  for(j in germ.lab){
    for(k in c(1:dim(lig.rec.soma.germ)[1])){
        lig <- as.character(lig.rec.soma.germ[k,1])
        rec <- as.character(lig.rec.soma.germ[k,2])
        int.pairs <- paste(lig,rec,sep = "-")
        cel.pairs <- paste(i,j,sep = "-")
        int.scores <- soma.mean[lig,i] * germ.mean[rec,j]
        int.cells.mat[int.pairs,cel.pairs] <- int.scores
    }
  }
}

```

```{r}
all.scores <- as.vector(int.cells.mat)
quant.95 <- quantile(all.scores, probs = 0.95)
int.cells.mat.trc <- int.cells.mat
int.cells.mat.trc[int.cells.mat<quant.95] = 0
```

```{r}
int.mean <- apply(int.cells.mat.trc,1,mean)
int.cells.mat.trc <- int.cells.mat.trc[int.mean!=0, ]
```

```{r}
Heatmap(int.cells.mat.trc, cluster_rows = FALSE, cluster_columns = FALSE)
```

```{r}
saveRDS(int.cells.mat, file = "~/storage/HumanMacaqueMouseCompAnaysis/interdata/ehvg.mou.int.cells.mat.rds")
saveRDS(int.cells.mat.trc, file = "~/storage/HumanMacaqueMouseCompAnaysis/interdata/ehvg.mou.int.cells.mat.trc.rds")
```

